---
title: "Flooded Land Inventory Estimate for Louisiana"
author: "J. Beaulieu"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
    toc: true
    toc_float: true
    fig_caption: yes
    depth: 2
    number_sections: true
    code_folding:  hide
editor_options: 
  chunk_output_type: console
---

## Setup

Load libraries and print session info to facilitate reproducability.
```{r setup, message=FALSE, warning=FALSE}

library(tidyverse)
library(sf)
library(USAboundaries)
library(lwgeom)
library(kableExtra)

# Print R session info
print(sessionInfo())
```


## Louisiana Flooded Lands
Flooded Lands in the US are derived from the [U.S. Army Corps of Engineers National Inventory of Dams](https://nid.sec.usace.army.mil/ords/f?p=105:1::::::), Navigable Waterways dataset maintained by the Department of Homeland security, and the [National Hydrography Data (NHD)](https://www.usgs.gov/core-science-systems/ngp/national-hydrography) maintained by the USGS.  Several screening tools were used to discriminate natural water bodies from Flooded Lands in these data sets.




```{r, message=FALSE, warning=FALSE, results="hide"}

# read data from local sources
inund <- sf::st_read(paste0("C:\\Users\\JBEAULIE\\Environmental Protection Agency (EPA)\\",
                            "SuRGE Survey of Reservoir Greenhouse gas Emissions - Documents\\",
                            "surgeDsn\\nationalFloodedLandsMapping\\Reservoir Data\\Data.gdb"),
                     layer = "InundationArea")
st_crs(inund) # 4269

nav <- sf::st_read(paste0("C:\\Users\\JBEAULIE\\Environmental Protection Agency (EPA)\\",
                            "SuRGE Survey of Reservoir Greenhouse gas Emissions - Documents\\",
                            "surgeDsn\\nationalFloodedLandsMapping\\Reservoir Data\\Data.gdb"),
                     layer = "Navigable_Waterways")
st_crs(nav) #4269

less8 <- sf::st_read(paste0("C:\\Users\\JBEAULIE\\Environmental Protection Agency (EPA)\\",
                            "SuRGE Survey of Reservoir Greenhouse gas Emissions - Documents\\",
                            "surgeDsn\\nationalFloodedLandsMapping\\Reservoir Data\\Data.gdb"),
                     layer = "Reservoirs_Less8ha") %>%
  st_transform(4269)

great8 <- sf::st_read(paste0("C:\\Users\\JBEAULIE\\Environmental Protection Agency (EPA)\\",
                            "SuRGE Survey of Reservoir Greenhouse gas Emissions - Documents\\",
                            "surgeDsn\\nationalFloodedLandsMapping\\Reservoir Data\\Data.gdb"),
                     layer = "Reservoirs_Greater8ha")  %>%
  st_transform(4269)


# get state boundary
la <- us_states() %>% # get states map
  filter(name == "Louisiana") %>%
  st_transform(4269) # 4269, as above

# clip waterbodies to state
la.inund <- sf::st_intersection(x = inund, y = la)
la.nav <- sf::st_intersection(x = nav, y = la)
la.less8 <- sf::st_intersection(x = less8, y = la)
la.great8 <- sf::st_intersection(x = great8, y = la)



ggplot() +
  geom_sf(data = la) +
  geom_sf(data = la.inund, color = "blue", fill = "blue") +
  geom_sf(data = la.nav, color = "blue", fill = "blue") +
  geom_sf(data = la.less8, color = "blue", fill = "blue") +
  geom_sf(data = la.great8, color = "blue", fill = "blue") +
  ggtitle("Louisiana Flooded Lands")


```


## GHG Emissions and Removals from Louisiana Flooded Lands

```{r}
# surface area and emissions
sa <- data.frame(type = c("Flooded Lands", "Ponds"),
                 area.m2 = c(st_area(la.inund) %>% sum() %>% as.numeric() +
                               st_area(la.nav) %>% sum() %>% as.numeric() +
                               st_area(la.great8) %>% sum() %>% as.numeric(), 
                             st_area(la.less8) %>% sum() %>% as.numeric())
) %>%
  mutate(area.ha = round(area.m2 / 10000, digits = 0), # convert to hectares
         ch4Emission = ifelse(type == "ponds",
                              area.ha * 183, # table 7.12
                              area.ha * 141.1), # table 7.9
         dnStream = ch4Emission * 0.09,
         totalCh4 = ch4Emission + dnStream) %>%
  select(-area.m2)
```
We assume that all Louisiana Flooded Lands are greater than 20 years old and thus classified as 'Flooded Lands Remaining Flooded Lands'.  Only CH~4~ emissions are inventoried for this managed lands category, whereas CO~2~ emissions are also inventoried for Flooded Lands less than 20 years old.  Nitrous oxide emissions are not accounted here, but are captured in other IPCC methodologies.

Methane emissions from flooded surfaces are estimated using IPCC default emission factors of 183 kg CH~4~ Ha^-1^ year^-1^ and 141.1 kg CH~4~ Ha^-1^ year^-1^ for ponds (<8 Ha) and larger Flooded Lands, respectively, in the tropical/moist climate zone.  Emissions of CH~4~ produced in Flooded Lands, but released downstream of the dam (downstream emissions), are calculated as 9% of the CH~4~ emitted from the Flooded Lands surface, per the IPCC default.  Total CH~4~ emissions from Louisiana's Flooded Lands are estimated as `r sum(sa$totalCh4)` kg CH~4~ year^-1^.

```{r}
sa %>% kbl(caption = "Surface area and annual CH4 emissions for Louisiana Flooded Lands",
           col.names = c("Flooded Lands Type", "Surface area (Ha)", "Surface emissions (kg CH4/year)", 
                         "Downstream emissions (kg CH4/year)", "Total emissions (kg CH4/year)"),
           format.args = list(big.mark = ",")) %>% 
  kable_classic()
```